manuscript_figures

### 2024-06-12 scripts/manuscript_figures.Rmd
updated figures and models

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tigris)
library(car)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(ggspatial)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)

@@

# Plot panel A zoommed out RI and sites in rectangles
panelA <- ggplot() +
  geom_sf(data = ri_state, fill = "lightgray") + # State boundaries
  geom_sf(data = site_rectangles, color = "red2", fill = NA, size = 2) + # Add rectangles around sites
  geom_text(data = labels_tot, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
  coord_sf(xlim = c(-72, -71.0), ylim = c(41.1, 42.1), expand = FALSE) + # Zoom level
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.8, "in"), pad_y = unit(0.4, "in"),
                         style = north_arrow_fancy_orienteering) +
    # Change x and y axes names
  labs(x = "Longitude", y = "Latitude") +
  theme_classic() # Remove axes and background

@@

# Define label data
labels <- data.frame(
  name = c("Fort\nWetherill"),
  lat = c(41.486),
  lon = c(-71.368)
)

# Plot panel B for fort wetherill transects
panelB <- ggplot() +
  geom_sf(data = ri_state, fill = "lightgray") + # State boundaries
  geom_point(data = transect_fw,
             aes(x = Start.Longitude, y = Start.Latitude),
             color = "black", fill = "red", size = 3, shape = 21) + # Add start points of transects
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.9, "in"), pad_y = unit(0.2, "in"),
                         style = north_arrow_fancy_orienteering) +
    geom_text(data = labels, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
  # Change x and y axes names
  labs(x = "Longitude", y = "Latitude") +
  theme_classic() + # Remove axes and background
  xlim(c(-71.45, -71.3)) + # Set longitude range
  ylim(c(41.42, 41.5))   # Set latitude range

@@

# Save the plot to a file
ggsave("output/figures/figure_1B.png", plot = panelB, width = 8, height = 6)

@@

# Adjust latitude values by subtracting 0.0011
transect_kb <- transect_kb %>%
  mutate(
    Start.Latitude = Start.Latitude - 0.0001
  )

# Plot panel C for kings beach transects
panelC <- ggplot() +
  geom_sf(data = ri_state, fill = "lightgray") + # State boundaries
  geom_point(data = transect_kb,
             aes(x = Start.Longitude, y = Start.Latitude),
             color = "black", fill = "red", size = 3, shape = 21) + # Add start points of transects
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.8, "in"), pad_y = unit(0.2, "in"),
                         style = north_arrow_fancy_orienteering) +
    geom_text(data = labels_kb, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
  # Change x and y axes names
  labs(x = "Longitude", y = "Latitude") +
  theme_classic() + # Remove axes and background
  xlim(c(-71.45, -71.3)) + # Set longitude range
  ylim(c(41.42, 41.5))   # Set latitude range

@@

# Combine the plots with patchwork
combined_plot <- (panelA | (panelB / panelC)) +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"))

# Save the combined plot
ggsave("output/figures/figure_1.png", combined_plot, width = 10, height = 6, units = "in", dpi = 300)

@@

# Create scatter plot
figure.2 <- ggplot(aggregated_data_cover, aes(x = mean_kelp_cover, y = mean_fish_biomass)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Kelp cover (%)", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Cover") +
  theme_classic();figure.2

@@

# Create scatter plot
figure.3 <- ggplot(aggregated.data.morph, aes(x = mean_kelp_biomass, y = mean_fish_biomass)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_biomass, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Kelp Biomass (g/m2)", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Biomass") +
  theme_classic();figure.3

@@

#delete unnecessary columns from the new joined data sheet 
all.params.data <- master.dat[, -c(1:7, 9:11, 14)]

#rename various columns in data frame
names(all.params.data)[1] <- "Kelp Cover (%)"
names(all.params.data)[2] <- "Fish Biomass (g/cm2)"
names(all.params.data)[3] <- "Rugosity"
names(all.params.data)[4] <- "Kelp Biomass (g/cm2)"

#reorder columns to show biomass by eaxchother
all.params.data <- all.params.data[, c(3, 1, 2, 4)]

@@

@@

ggsave(filename = "output/figures/figure_5.png", device = "png", width = 10, height = 10)
```

>>>>

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tigris)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)

@@

# Plot panel A zoommed out RI and sites in rectangles
panelA <- ggplot() +
  geom_sf(data = ri_state, fill = "lightgray") + # State boundaries
  geom_sf(data = site_rectangles, color = "red2", fill = NA, size = 2) + # Add rectangles around sites
  geom_text(data = labels_tot, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
  coord_sf(xlim = c(-72, -71.0), ylim = c(41.1, 42.1), expand = FALSE) + # Zoom level
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(1.0, "in"), pad_y = unit(0.2, "in"),
                         style = north_arrow_fancy_orienteering) +
    # Change x and y axes names
  labs(x = "Longitude", y = "Latitude") +
  theme_classic() # Remove axes and background

@@

# Define label data
labels <- data.frame(
  name = c("Fort\nWetherill"),
  lat = c(41.486),
  lon = c(-71.37)
)

# Plot panel B for fort wetherill transects
panelB <- ggplot() +
  geom_sf(data = ri_state, fill = "lightgray") + # State boundaries
  geom_point(data = transect_fw,
             aes(x = Start.Longitude, y = Start.Latitude),
             color = "black", fill = "red", size = 3, shape = 21) + # Add start points of transects
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.5, "in"), pad_y = unit(0.2, "in"),
                         style = north_arrow_fancy_orienteering) +
    geom_text(data = labels, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
  # Change x and y axes names
  labs(x = "Longitude", y = "Latitude") +
  theme_classic() + # Remove axes and background
  xlim(c(-71.40, -71.35)) + # Set longitude range
  ylim(c(41.46, 41.5))   # Set latitude range

@@

# Save the plot to a file
ggsave("output/figures/figure_1B.png", plot = panelB, width = 8, height = 8)

@@

# Adjust latitude values by subtracting 0.0011
transect_kb <- transect_kb %>%
  mutate(
    Start.Latitude = Start.Latitude - 0.0003
  )

# Plot panel C for kings beach transects
panelC <- ggplot() +
  geom_sf(data = ri_state, fill = "lightgray") + # State boundaries
  geom_point(data = transect_kb,
             aes(x = Start.Longitude, y = Start.Latitude),
             color = "black", fill = "red", size = 3, shape = 21) + # Add start points of transects
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.6, "in"), pad_y = unit(0.2, "in"),
                         style = north_arrow_fancy_orienteering) +
    geom_text(data = labels_kb, aes(x = lon, y = lat, label = name), size = 5, color = "black") + # Add labels
  # Change x and y axes names
  labs(x = "Longitude", y = "Latitude") +
  theme_classic() + # Remove axes and background
  xlim(c(-71.40, -71.3)) + # Set longitude range
  ylim(c(41.44, 41.5))   # Set latitude range

@@

# Combine the plots with patchwork
combined_plot <- (panelA | (panelB / panelC)) +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));combined_plot

# Save the combined plot
ggsave("output/figures/figure_1.png", combined_plot, width = 12, height = 8, units = "in", dpi = 300)

@@

# Create scatter plot
figure.2 <- ggplot(aggregated_data_cover, aes(x = log(mean_kelp_cover), y = log(mean_fish_biomass))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Log Kelp cover (%)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Cover") +
  theme_classic();figure.2

@@

# Create scatter plot
figure.3 <- ggplot(aggregated.data.morph, aes(x = log(mean_kelp_biomass), y = log(mean_fish_biomass))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_biomass, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Log Kelp Biomass (g/m2)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Biomass") +
  theme_classic();figure.3

@@

#select necessary columns
all.params.data <- master.dat %>% select("YEAR", "SITE", "COVER", "Fish.Biomass..g.m2.", "mean_rugosity", "average_kelp_biomass")

#rename various columns in data frame
names(all.params.data)[3] <- "Kelp Cover (%)"
names(all.params.data)[4] <- "Fish Biomass (g/cm2)"
names(all.params.data)[5] <- "Rugosity"
names(all.params.data)[6] <- "Kelp Biomass (g/cm2)"

#reorder columns to show biomass by eachother
all.params.data <- all.params.data[, c(1,2, 4, 6, 5, 3)]

@@

# Define the custom apply_rules function
apply_rules <- function(p_values, rules = c(0.001, 0.01, 0.05), symbols = c("***", "**", "*")) {
  result <- rep("", length(p_values))
  for (i in seq_along(rules)) {
    result[p_values < rules[i]] <- symbols[i]
  }
  return(result)
}

@@

ggsave(filename = "output/figures/correlation_analysis_figure.png", device = "png", width = 10, height = 10)

# Figure 5: Multiple regression between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
# combine all dataframes
```{r}
# Rename columns to remove spaces and special characters
names(all.params.data) <- c("YEAR", "SITE", "Fish_Biomass", "Rugosity", "Kelp_Cover", "Kelp_Biomass")

# was getting an error: qr.default(X, tol = tol, LAPACK = FALSE) when running model below which indicates that there are still problematic values (NA, NaN, or Inf) in the data used for the model fitting

# Check for zero or negative values
zero_or_negative <- sapply(all.params.data[, c("Fish_Biomass", "Rugosity", "Kelp_Biomass", "Kelp_Cover")], function(x) sum(x <= 0))
zero_or_negative

# Remove rows with zero or negative values in relevant columns
all.params.data <- all.params.data[all.params.data$Fish_Biomass > 0 & 
                                   all.params.data$Rugosity > 0 & 
                                   all.params.data$Kelp_Biomass > 0 & 
                                   all.params.data$Kelp_Cover > 0, ]

# Fit a multiple regression model
fit1 <- lm(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover), data = all.params.data)

# Diagnostic plots for the regression model
par(mfrow=c(2,2))
plot(fit1)

# Produce added multivariable plots
av_plots <- avPlots(fit1)

#The x-axis displays a single predictor variable and the y-axis displays the response variable.
#The blue line shows the association between the predictor variable and the response variable, while holding the value of all other predictor variables constant.
#The points that are labelled in each plot represent the 2 observations with the largest residuals and the 2 observations with the largest partial leverage.

#view results of model
summary_text1 <- capture.output(summary(fit1))
writeLines(summary_text1, "output/figures/summary_original_model.txt")

# Fit a multiple regression model with site and year as interactions
fit2 <- lm(Fish_Biomass ~ Rugosity * Kelp_Biomass * Kelp_Cover * SITE * YEAR, data = all.params.data)

# Diagnostic plots for the regression model
par(mfrow=c(2,2))
plot(fit2)

#view results of model
summary_text2 <- capture.output(summary(fit2))
writeLines(summary_text2, "output/figures/summary_model_site_year_fixed.txt")

# Try another type of multiple regession plot
```{r}
#fit multiple linear regression model WITH SITE AND YEAR
fit3 <- lm(Fish_Biomass ~ SITE*YEAR, data = all.params.data)

#view results of model
summary_text3 <- capture.output(summary(fit3))
writeLines(summary_text3, "output/figures/summary_model_site_year_fixed_fishbiomass.txt")

# Fit the mixed-effects model
fit4 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
              (1 | SITE) + (1 | YEAR), data = all.params.data)

#view results of model
summary_text4 <- capture.output(summary(fit4))
writeLines(summary_text4, "output/figures/summary_model_site_year_random.txt")

# site and year no significant interaction or significant main effects

# run with just site as random effect
# Fit the mixed-effects model
fit5 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
              (1 | SITE), data = all.params.data)

#view results of model
summary_text5 <- capture.output(summary(fit5))
writeLines(summary_text5, "output/figures/summary_model_site_random.txt")

# site and year no significant main effects

# run with just year as random effect
# Fit the mixed-effects model
fit6 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
              (1 | YEAR), data = all.params.data)

#view results of model
summary_text6 <- capture.output(summary(fit6))
writeLines(summary_text6, "output/figures/summary_model_year_random.txt")

#year is significantly different for kelp cover
